---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tiltIndicator

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/2DegreesInvesting/tiltIndicator/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/2DegreesInvesting/tiltIndicator/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/2DegreesInvesting/tiltIndicator/branch/main/graph/badge.svg)](https://app.codecov.io/gh/2DegreesInvesting/tiltIndicator?branch=main)
<!-- badges: end -->

The goal of tiltIndicator is to help you develop each TILT indicator.

This repository hosts only public code and may only show only fake data.

## Installation

You can install the development version of tiltIndicator from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("2DegreesInvesting/tiltIndicator")
```

## Example

```{r}
library(tiltIndicator)
```

We'll use datasets that come with the tiltIndicator package.

```{r}
pstr_companies

pstr_ep_weo

pstr_weo_2022
```

The `pstr_*()` functions help you work with the product-sector-transition-risk
indicator.

```{r}
with_scores <- pstr_companies |>
  pstr_add_reductions(pstr_ep_weo, pstr_weo_2022) |>
  pstr_add_transition_risk() |>
  pstr_aggregate_scores(pstr_companies)

with_scores
```
