---
title: "Handling a long runtime"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This article shows how to calculate an indicator when it takes too long to run.
It shows an approach that shows your progress, avoids running the same thing
twice, and allows you to restart from where you left if something fails or you
need to interrupt.

### Setup

We'll need a number of packages. 

```{r}
library(dplyr, warn.conflicts = FALSE)
library(readr, warn.conflicts = FALSE)
library(progress)
library(tiltIndicator)
library(fs)

packageVersion("tiltIndicator")
```

Here I use example data from the tiltIndicator package.

```{r}
# TODO: Replace with your real `companies`
real_companies <- companies
# TODO: Replace your real `inputs` for ICTR
real_co2 <- products
```

This example calculates the indicator "PCTR". Adapt it as necessary.

```{r}
# TODO: Replace with the literal string "ictr" for ICTR
indicator <- "pctr"

# Create a folder to store the results of each indicator
# TODO: Replace `tempdir()` with a permanent folder, e.g. ~/Downloads
parent <- tempdir()
dir_create(path(parent, indicator))
```

Split the data by company. 

```{r}
companies_list <- split(real_companies, real_companies$company_id)
```

Calculate the indicator for each company at a time, saving the result to a .csv
file, and skipping companies that are already done.

```{r}
# Setup a progress bar (pb)
pb <- progress_bar$new(total = length(companies_list))
for (i in seq_along(companies_list)) {
  pb$tick()
  
  # Skip if run previously
  companies_id <- names(companies_list[i])
  file <- path(parent, indicator, paste0(companies_id, ".csv"))
  if (exists(file)) next()
  
  # Run
  result <- xctr(companies_list[[i]], real_co2)
  
  # Save into an indicator-specific folder, i.e. either ictr/ or pctr/
  write_csv(result, file)
}
```

When you're done (or before) you can read all the .csv files into a single
dataset at once.

```{r}
combined_results <- read_csv(dir_ls(path(parent, indicator)))
combined_results
```

You may want to run this on as a [background
job](https://docs.posit.co/ide/user/ide/guide/tools/jobs.html) so you can use
your R session for something else while the process runs on the background. Or
better, you may run it on a remote server so you can rent a massive computer for
a short time, and use your time and laptop for other things.
