---
title: "PCTR questions"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of this article is to communicate what I found while characterizing
PCTR data and tests. 

Here I show code to ask questions that might help us reduce the size of the
example data, define the code's intended behavior, and maybe find bugs and
potential improvements. I start with an overview, then explore required columns
and rows.

## Overview

```{r setup}
library(dplyr, warn.conflicts = TRUE)
library(tiltIndicator)

packageVersion("tiltIndicator")
```

You can find all PCTR data and functions in RStudio with auto-complete.

<img src=https://user-images.githubusercontent.com/5856545/231594835-7ff10ec8-5e91-4362-a5c2-23718c8afc34.png width=800>

You can use it all in a single pipe.

```{r}
out <- pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies)

out
```

## 1. Required columns

It's best to [minimize the size of the inputs and
outputs](https://gist.github.com/maurolepore/17b468dd8c3a603c668220563754ad7f#minimize-the-size-of-the-inputs-and-ouptuts),
so that the data is easier to understand and easier to work with. Smaller data
fits better on screen, uses less disk space, and takes less time to compute on.

Characterization tests show that we may be able to remove columns from both PCTR
datasets, and from the output of `pctr_score_activities()`. Here is the same
pipe I showed above but this time with the proposed `*_new` version of data and
code.

```{r}
#| echo=FALSE
pctr_ecoinvent_co2_new <- pctr_ecoinvent_co2 |>
  select(all_of(tiltIndicator:::pctr_ecoinvent_co2_crucial()))

pctr_companies_new <- pctr_companies |>
  select(all_of(tiltIndicator:::pctr_companies_crucial()))

pctr_score_activities_new <- function(...) {
  pctr_score_activities(...) |>
    select(all_of(tiltIndicator:::pctr_score_companies_crucial()))
}
```

```{r}
out_new <- pctr_ecoinvent_co2_new |>
  pctr_score_activities_new() |>
  pctr_score_companies(pctr_companies_new)

out_new
```

Compare

```{r}
pctr_ecoinvent_co2 |> glimpse()
pctr_ecoinvent_co2_new |> glimpse()

pctr_companies |> glimpse()
pctr_companies_new |> glimpse()

pctr_ecoinvent_co2_new |>
  pctr_score_activities() |>
  glimpse()
pctr_ecoinvent_co2_new |>
  pctr_score_activities_new() |>
  glimpse()

# Same despite `out_new` results from smaller data
out |> glimpse()
out_new |> glimpse()
```

Questions:

1.1. Is it OK to remove the proposed columns?

1.2. Is there any additional column we could remove from `out_new`?

## 2. Required rows

In addition to removing needless columns, we may also be able to remove needless
rows, or restructure the data to make it easier to understand and work with.

`pctr_score_activities()` has a clear 1 to 1 relationship between the number of
rows of the input and output. If it gets 1 row it returns 1 row.

```{r}
pctr_ecoinvent_co2 |>
  slice(1) |>
  pctr_score_activities()

pctr_ecoinvent_co2 |>
  slice(1:2) |>
  pctr_score_activities()
```

`pctr_score_copmanies()` returns three rows for every input company.

```{r}
#| error=TRUE
first1_company <- pctr_companies |> filter(company_id %in% unique(company_id)[1])
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(first1_company) |>
  dim()

first2_company <- pctr_companies |> filter(company_id %in% unique(company_id)[1:2])
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(first2_company) |>
  dim()

first3_company <- pctr_companies |> filter(company_id %in% unique(company_id)[1:3])
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(first3_company) |>
  dim()
```

The number of rows in the output doesn't depend on the number of rows in the
`pctr_ecoinvent_co2` dataset.

```{r}
#| error=TRUE
first1_company <- pctr_companies |> filter(company_id %in% unique(company_id)[1])
pctr_ecoinvent_co2 |>
  slice(1) |>
  pctr_score_activities() |>
  pctr_score_companies(first1_company) |>
  dim()
pctr_ecoinvent_co2 |>
  slice(1:2) |>
  pctr_score_activities() |>
  pctr_score_companies(first1_company) |>
  dim()
pctr_ecoinvent_co2 |>
  slice(1:10) |>
  pctr_score_activities() |>
  pctr_score_companies(first1_company) |>
  dim()
```

Questions:

2.1. Why is `pctr_ecoinvent_co2` 25 rows long?
2.2. Can we reduce the size of `pctr_ecoinvent_co2` to, say, 5 rows?
