---
title: "PCTR questions"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of this article is to communicate what I found while characterizing
PCTR data and tests. 

Here I show code and ask questions that migh help me reduce
the size of the example data, define the code's intended behaviour, and maybe
find bugs and potential improvements. I start with an overview, then explore required columns and rows.

## Overview

```{r setup}
library(tidyverse)
devtools::load_all()

packageVersion("tiltIndicator")
```

You can find all PCTR data and functions in RStudio with auto-complete.

<img src=https://user-images.githubusercontent.com/5856545/231594835-7ff10ec8-5e91-4362-a5c2-23718c8afc34.png width=800>

You can use it all in a single pipe.

```{r}
out <- pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies)

out
```

## 1. Required columns

It's best to [miminize the size of the inputs and
outputs](https://gist.github.com/maurolepore/17b468dd8c3a603c668220563754ad7f#minimize-the-size-of-the-inputs-and-ouptuts),
so the data is easier to understant and easier to work with. Smaller data fits
better on screen, takes up less spaces and less time to compute on.

Characterization tests show we can remove columns from both datasets, and from
the output of `pctr_score_activities()`.

```{r}
#| echo=FALSE

pctr_ecoinvent_co2_new <- pctr_ecoinvent_co2 |>
  select(all_of(pctr_ecoinvent_co2_crucial()))

pctr_companies_new <- pctr_companies |>
  select(all_of(pctr_companies_crucial()))

pctr_score_activities_new <- function(...) {
  pctr_score_activities(...) |>
    select(all_of(pctr_score_companies_crucial()))
}
```

```{r}
out_new <- pctr_ecoinvent_co2_new |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies_new)

out_new
```

Compare

```{r}
pctr_ecoinvent_co2 |> glimpse()

pctr_ecoinvent_co2_new |> glimpse()

pctr_companies |> glimpse()

pctr_companies_new |> glimpse()

out |> glimpse()

out_new |> glimpse()
```

Questions:

1.1. Is it OK to remove the proposed columns?

1.2. Can we also remove any column from the final output (see `out2`)?

## 2. Required rows

`pctr_score_activities()` has a clear, 1 to 1 realationship between the number
of rows of the input and output. If it gets 1 row it returns 1 row.

```{r}
pctr_ecoinvent_co2 |>
  slice(1) |>
  pctr_score_activities()

pctr_ecoinvent_co2 |>
  slice(1:2) |>
  pctr_score_activities()
```

`pctr_score_copmanies()` is puzzling. If I change the number of rows of
`pctr_ecoinvent_co2` I can't predict if the code will work of fail.

```{r}
#| error=TRUE
# Fails
pctr_ecoinvent_co2 |>
  slice(1) |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies)

pctr_ecoinvent_co2 |>
  slice(1:15) |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies)

pctr_ecoinvent_co2 |>
  slice(4:15) |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies)

# Fails
pctr_ecoinvent_co2 |>
  slice(5:15) |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies)
```

The same unpredictable success happens if I change the number or rows in the
`pctr_companies` dataset.

```{r}
#| error=TRUE
# Fails
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies |> slice(1))

pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies |> slice(1:10))

pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies |> slice(4:10))

# Fails
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies |> slice(5:15))
```


Questions:

Well designed code is easy to understand and use, so this suggests I need to
improve it. What might be the problem?

2.1. The way I divided the code across the two functions isn't the best.

2.2. There is a bug.

2.3. The logic is clear but I don't understand it.
