---
title: "PCTR questions"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here I ask the questions I have when I characterize the PCTR data and code.

```{r setup}
library(dplyr, warn.conflicts = TRUE)
library(tiltIndicator)

packageVersion("tiltIndicator")
```

## 1. Joining by redundant columns

Questions:

* 1.1. We call `*_join()` with `by = c("activity_product_uuid", "ei_activity",
"unit")` and Tilman suggested we might remove `"ei_activity"` and `"unit"`. 
Do we expect them to be always redundant with the groups defined by
`"activity_product_uuid"` alone? Or do you mean they are redundant in this
specific toy dataset?

## 2. Edge case with too few co2 data

`pctr_score_copmanies()` returns 3 rows for each company (regardless the number
of rows per company). This seems expected.

```{r}
#| error=TRUE
one_company <- pctr_companies |> filter(company_id %in% first(company_id))
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(one_company)

another_company <- pctr_companies |> filter(company_id %in% last(company_id))
two_companies <- bind_rows(one_company, another_company)
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(two_companies)
```

But the number number of rows in the co2 data can be such that the output seems
useless and you get no warning.

```{r}
#| error=TRUE
too_few_co2_rows <- pctr_ecoinvent_co2 |> slice(1)
too_few_co2_rows

too_few_co2_rows |>
  pctr_score_activities() |>
  pctr_score_companies(one_company)

too_few_co2_rows |>
  pctr_score_activities() |>
  pctr_score_companies(two_companies)
```

Questions:

* 2.1. Is this output useful?
* 2.2. Should we allow this output with a warning or throw an error and stop?
