---
title: "PCTR questions"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of this article is to communicate what I found while characterizing
PCTR data and tests. 

Here I show code to ask questions that might help us reduce the size of the
example data, define the code's intended behavior, and maybe find bugs and
potential improvements. I start with an overview, then explore required columns
and rows.

## Overview

```{r setup}
library(dplyr, warn.conflicts = TRUE)
library(tiltIndicator)

packageVersion("tiltIndicator")
```

You can find all PCTR data and functions in RStudio with auto-complete.

<img src=https://user-images.githubusercontent.com/5856545/231594835-7ff10ec8-5e91-4362-a5c2-23718c8afc34.png width=800>

You can use it all in a single pipe.

```{r}
out <- pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies)

out
```

## 1. Required columns

It's best to [miminize the size of the inputs and
outputs](https://gist.github.com/maurolepore/17b468dd8c3a603c668220563754ad7f#minimize-the-size-of-the-inputs-and-ouptuts),
so that the data is easier to understand and easier to work with. Smaller data
fits better on screen, uses less disk space, and takes less time to compute on.

Characterization tests show that we may be able to remove columns from both PCTR
datasets, and from the output of `pctr_score_activities()`. Here is the same
pipe I showed above but this time with the proposed `*_new` version of data and
code.

```{r}
#| echo=FALSE
pctr_ecoinvent_co2_new <- pctr_ecoinvent_co2 |>
  select(all_of(tiltIndicator:::pctr_ecoinvent_co2_crucial()))

pctr_companies_new <- pctr_companies |>
  select(all_of(tiltIndicator:::pctr_companies_crucial()))

pctr_score_activities_new <- function(...) {
  pctr_score_activities(...) |>
    select(all_of(tiltIndicator:::pctr_score_companies_crucial()))
}
```

```{r}
out_new <- pctr_ecoinvent_co2_new |>
  pctr_score_activities_new() |>
  pctr_score_companies(pctr_companies_new)

out_new
```

Compare

```{r}
pctr_ecoinvent_co2 |> glimpse()

pctr_ecoinvent_co2_new |> glimpse()

pctr_companies |> glimpse()

pctr_companies_new |> glimpse()

out |> glimpse()

out_new |> glimpse()
```

Questions:

1.1. Is it OK to remove the proposed columns?

1.2. Is there any additional column we could remove from `out_new`?

## 2. Required rows

In addition to removing needless columns, we may also be able to remove needless
rows, or restructure the data to make it easier to understand and work with. But
I'm struggling to understand how to do that. Soon I'll learn more from running
the code in debugger mode, but some of my questions may already be clear to
Tilman, so I share this maybe a bit before I'm fully done with my own thinking.

`pctr_score_activities()` seems straightforward. It has a clear 1 to 1
relationship between the number of rows of the input and output. If it gets 1
row it returns 1 row.

```{r}
pctr_ecoinvent_co2 |>
  slice(1) |>
  pctr_score_activities()

pctr_ecoinvent_co2 |>
  slice(1:2) |>
  pctr_score_activities()
```

But `pctr_score_copmanies()` is puzzling. If I change the number of rows of
`pctr_ecoinvent_co2` I can't predict if the code will work of fail.

```{r}
#| error=TRUE
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies) |>
  dim()

# Fails
pctr_ecoinvent_co2 |>
  slice(1) |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies) |>
  dim()

pctr_ecoinvent_co2 |>
  slice(1:15) |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies) |>
  dim()

pctr_ecoinvent_co2 |>
  slice(4:15) |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies) |>
  dim()

# Fails
pctr_ecoinvent_co2 |>
  slice(5:15) |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies) |>
  dim()
```

Similarly puzzling is what happens if I change the number or rows in the
`pctr_companies` dataset. The oddest behavior is that it seems that all
companies are required rather than being a parameter we're free to change.

```{r}
#| error=TRUE
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(pctr_companies) |>
  dim()

# Fails
first_row <- pctr_companies |> slice(1)
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(first_row) |>
  dim()

# Fails
first_company <- pctr_companies |> filter(company_id %in% first(company_id))
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(first_company) |>
  dim()

first_row_of_each_company <- pctr_companies |>
  group_by(company_id) |>
  slice(1) |>
  ungroup()
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(first_row_of_each_company) |>
  dim()

# Fails (oddest) Why do we need every one of these companies?
first_row_of_two_out_of_three_companies <- pctr_companies |>
  group_by(company_id) |>
  slice(1) |>
  ungroup() |>
  slice(1:2)
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(first_row_of_two_out_of_three_companies) |>
  dim()

first_two_rows_of_each_company <- pctr_companies |>
  group_by(company_id) |>
  slice(1:2) |>
  ungroup()
pctr_ecoinvent_co2 |>
  pctr_score_activities() |>
  pctr_score_companies(first_two_rows_of_each_company) |>
  dim()
```

Questions:

What might be the problem?

2.1. The way I divided the code across the two functions isn't the best.

2.2. There is a bug.

2.3. The problem is clear to Tilman but I still don't understand it.
