---
title: "The tilt indicators"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This article helps you get started with the tiltIndicator package. Currently it
supports two families of indicators: "emissions" and "sector". 

In the "emissions" family, you have two indicators -- "emissions profile" and
"emissions profile upstream". They are both calculated in exactly the same way.
The difference is not in the process but in the dataset holding CO2 data: Either
it has products or upstream-products (a.k.a. inputs), respectively.

In the "sector" family, you also have two indicators: "Sector profile" and
"sector profile upstream". Also they are calculated in the same way, but "sector
profile upstream" takes an additional argument for upstream-products (a.k.a.
inputs).

## Emissions profile

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=tiltIndicator:::extdata_path("child/intro-emissions_profile.Rmd")}
```

### Example

```{r child=tiltIndicator:::template_to_rmd("example-emissions_profile.R")}
```

## Emissions profile upstream

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=tiltIndicator:::extdata_path("child/intro-emissions_profile_upstream.Rmd")}
```

### Example

```{r child=tiltIndicator:::template_to_rmd("example-emissions_profile_upstream.R")}
```

## Sector profile

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=tiltIndicator:::extdata_path("child/intro-sector_profile.Rmd")}
```

### Example

```{r child=tiltIndicator:::template_to_rmd("example-sector_profile.R")}
```

## Sector profile upstream

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=tiltIndicator:::extdata_path("child/intro-sector_profile_upstream.Rmd")}
```

### Example

```{r child=tiltIndicator:::template_to_rmd("example-sector_profile_upstream.R")}
```

## Thresholds

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=tiltIndicator:::extdata_path("child/thresholds.Rmd")}
```

## Enriching the output

All profile functions pass any `*rowid` column from the input to the output at
product level. This allows you to enrich the default output.

```{r}
library(dplyr, warn.conflicts = FALSE)

products_with_rowid <- products |> tibble::rowid_to_column("products_rowid")
products_with_rowid

result_with_rowid <- companies |> emissions_profile(products_with_rowid)

rich_products_result <- result_with_rowid |>
  unnest_product() |>
  left_join(products_with_rowid)

rich_products_result |> relocate(matches("sector"))
```

