---
title: "The tilt indicators"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This article introduces the tilt indicators. First we introduce the tilt
indicators in general, then explain each indicator's details and show the code
you can use to calculate it.

We'll use general purpose tools from the dplyr package, and the experimental [tiltIndicator](https://2degreesinvesting.github.io/tiltIndicator/index.html) package.

```{r setup}
library(dplyr, warn.conflicts = FALSE)
library(tiltIndicator)
```

The tiltIndicator package supports two indicators: Input Carbon Transition Risk
(ICTR) and Product Carbon Transition Risk (PCTR). They are both calculated in
exactly the same way. The difference is not the process but the kind of `co2`
data: either comes from inputs or products.

## Input carbon transition risk (ICTR)

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=extdata_path("child/intro-ictr.Rmd")}
```

## Product carbon transition risk (PCTR)

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=extdata_path("child/intro-pctr.Rmd")}
```

## Example

Use `xctr*()` to calculate both ICTR and PCTR. Their difference is in the `co2`
dataset. Here we'll use example datasets that come with the tiltIndicator
package: `companies`, `inputs`, and `products`.

```{r}
companies

# ICTR
inputs

companies |>
  xctr_at_product_level(co2 = inputs)

companies |>
  xctr_at_product_level(co2 = inputs) |>
  xctr_at_company_level()

# Same
companies |> xctr(co2 = inputs)

# PCTR
products

companies |>
  xctr_at_product_level(co2 = products)

companies |>
  xctr_at_product_level(co2 = products) |>
  xctr_at_company_level()

# Same
companies |> xctr(co2 = products)
```
