---
title: "The tilt indicators"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This article helps you get started with the tiltIndicator package. Currently it
supports two families of indicators: "emissions" and "sector". In the
"emissions" family, you have two indicators: "Emissions profile" and "emissions
profile upstream". They are both calculated in exactly the same way. The
difference is not the process but the kind of `co2` data: Either comes from
products or upstream products.

In the "sector" family, you have two indicators: "Sector profile" and "sector
profile upstream". They are both calculated in the same way, but "sector profile
upstream" takes an additional argument for upstream products.

We'll use general purpose tools from the dplyr package, and the experimental
[tiltIndicator](https://2degreesinvesting.github.io/tiltIndicator/index.html)
[tiltToyData](https://2degreesinvesting.github.io/tiltIndicator/index.html)
packages.

```{r setup}
library(tiltIndicator)
library(tiltToyData)
library(dplyr, warn.conflicts = FALSE)
library(readr, warn.conflicts = FALSE)
library(tibble)

options(readr.show_col_types = FALSE)
```

## Emissions profile

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=extdata_path("child/intro-emissions-profile.Rmd")}
```

## Emissions profile upstream

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=extdata_path("child/intro-emissions-profile-upstream.Rmd")}
```

## Thresholds

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=extdata_path("child/thresholds.Rmd")}
```

## Example

Use `emissions_profile()` and `emissions_profile_upstream()` with example
datasets that come with the tiltToyData package.

```{r}
toy_files()

companies <- read_csv(toy_emissions_profile_any_companies())
products <- read_csv(toy_emissions_profile_products())
products_upstream <- read_csv(toy_emissions_profile_upstream_products())

result <- emissions_profile(companies, products)
result

result |> unnest_product()

result |> unnest_company()

result_upstream <- emissions_profile_upstream(companies, products_upstream)
result_upstream

result_upstream |> unnest_product()

result_upstream |> unnest_company()
```

### Enriching the output

All profile functions pass any `*rowid` column from the input to the output at
product level. This allows you to enrich the default output.

```{r}
products_with_rowid <- products |> tibble::rowid_to_column("products_rowid")
products_with_rowid

result_with_rowid <- companies |> emissions_profile(products_with_rowid)

result_rich <- result_with_rowid |>
  unnest_product() |> 
  left_join(products_with_rowid)

result_rich |> relocate(matches("sector"))
```

