---
title: "The tilt indicators"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This article helps you get started with the tiltIndicator package. Currently it
supports two families of indicators: XCTR and XSTR. In the "emissions" family,
you have two indicators: "Emissions profile" and "emissions profile upstream".
They are both calculated in exactly the same way. The difference is not the
process but the kind of `co2` data: Either comes from products or upstream
products.

In the "sector" family, you have two indicators: "Sector profile" and "sector
profile upstream". They are both calculated in the same way, but "sector profile
upstream" takes an additional argument for upstream products.

You can find the main functions of the families
[here](https://2degreesinvesting.github.io/tiltIndicator/reference/index.html#main-functions-and-datasets).

We'll use general purpose tools from the dplyr package, and the experimental
[tiltIndicator](https://2degreesinvesting.github.io/tiltIndicator/index.html)
[tiltToyData](https://2degreesinvesting.github.io/tiltIndicator/index.html)
packages.

```{r setup}
library(tiltIndicator)
library(tiltToyData)
library(dplyr, warn.conflicts = FALSE)
library(readr, warn.conflicts = FALSE)

options(readr.show_col_types = FALSE)
```

## Emissions profile

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=extdata_path("child/intro-pctr.Rmd")}
```

## Emissions profile upstream

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=extdata_path("child/intro-ictr.Rmd")}
```

## Thresholds

<!-- WARNING: Do not write here. Instead write in the `child` file -->

```{r child=extdata_path("child/thresholds.Rmd")}
```

## Example

Use `emissions_profile()` and `emissions_profile_upstream()` with example
datasets that come with the tiltToyData package.

```{r}
toy_files()
companies <- read_csv(toy_path("emissions_profile_any_companies.csv.gz"))
products <- read_csv(toy_path("emissions_profile_products.csv.gz"))
products_upstream <- read_csv(toy_path("emissions_profile_upstream_products.csv.gz"))

result <- emissions_profile(companies, products)
result

result |> unnest_product()

result |> unnest_company()

result_upstream <- emissions_profile_upstream(companies, products_upstream)
result_upstream

result_upstream |> unnest_product()

result_upstream |> unnest_company()
```

